---
title: "class14"
author: "Nattiwong"
date: "3/4/2022"
output: github_document
---


```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
```

```{r message=FALSE}
library(dplyr)
#install.packages("skimr")
library(skimr)
#install.packages("lubridate")
library(lubridate)
#install.packages("zipcodeR")
library(zipcodeR)
library(tidyverse)
```

```{r}
skimr::skim(vax)
```

Q1. What column details the total number of people fully vaccinated? 
```{r}
head(vax$persons_fully_vaccinated)
```

Q2. What column details the Zip code tabulation area? 

```{r}
head(vax$zip_code_tabulation_area)
```

Q3. What is the earliest date in this dataset? 

```{r}
vax[1,1]
```

Q4. What is the latest date in this dataset? 

```{r}
vax[nrow(vax),1]
```


Q5. How many numeric columns are in this dataset? 
10 columns

Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```

Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

```{r}
round((sum(is.na(vax$persons_fully_vaccinated))/length(vax$persons_fully_vaccinated))*100,2)
```

Q8. [Optional]: Why might this data be missing?


```{r}
vax$as_of_date <- ymd(vax$as_of_date)
today() - vax$as_of_date[1]
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

Q9. How many days have passed since the last update of the dataset? 
```{r}
today() - vax$as_of_date[nrow(vax)]
```

Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 
```{r}
length(unique(ymd(vax$as_of_date)))
```

Overwrite the as_of_date column by the output of ymd() function
```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```


```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```


#focus on the San Diego area

```{r}
head(vax[vax$county == "San Diego",])
```

Select some specific row by filter function in dplyr package

```{r}
sd <- filter(vax, county == "San Diego")
#vax %>% filter(country=="San Diego")
```


```{r}
head(sd)
nrow(sd)
```
 
 Q11. How many zip codes are there in San Diego country?
```{r}
length(unique(sd$zip_code_tabulation_area))
length(table(sd$zip_code_tabulation_area))
```

```{r}
vax %>%
    filter(county == "San Diego") %>%
    select(zip_code_tabulation_area) %>%
    unique %>%
    nrow()
```

Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset? 
```{r}
sd %>% 
    filter(age12_plus_population==max(sd$age12_plus_population)) %>%
    select(zip_code_tabulation_area) %>%
    unique()
```

```{r}
inds <- order(sd$age12_plus_population, decreasing = TRUE)
sd[inds[1],]
```
```{r}
ind <- which.max(sd$age12_plus_population)
sd$zip_code_tabulation_area[ind]
```

```{r}
arrange(sd, -age12_plus_population)[1,]
```

Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-02-22”? 

```{r}
sd.now <- sd %>% 
    filter(as_of_date==ymd("2022-02-22"))
mean(sd.now$percent_of_population_fully_vaccinated, na.rm = TRUE)

sd %>% 
    filter(as_of_date==ymd("2022-02-22")) %>%
    select(percent_of_population_fully_vaccinated) %>%
    colMeans(na.rm=T) #

```

Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-02-22”?

```{r}
hist(sd.now$percent_of_population_fully_vaccinated, 
     xlab="percent",
     ylab="frequency",
     main="Histogram of Vaccination Rate across San Diego County",
     sub="as of 2022-02-22")
```

```{r}
ggplot(sd.now) +
    aes(percent_of_population_fully_vaccinated) +
    geom_histogram() +
    geom_rug()
```

```{r}
ucsd <- filter(sd, sd$zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```
Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
ucsd.plot <- ggplot(ucsd) +
  aes(x=as_of_date,
      y=percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```

Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
lj <- filter(sd.now, sd.now$zip_code_tabulation_area=="92037")
lj$age5_plus_population
```

```{r}
vax.36 <- filter(vax, age5_plus_population > lj$age5_plus_population &
                as_of_date == "2022-02-22")
mean(vax.36$percent_of_population_fully_vaccinated)
```

```{r}
ucsd.plot +
    geom_hline(yintercept = mean(vax.36$percent_of_population_fully_vaccinated),
               color="red",
               linetype=2)
```

Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```

Q18. Using ggplot generate a histogram of this data.
```{r}
ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated) +
  geom_histogram() +
  labs(x="Percent Vaccinated", y="Count") +
    xlim(0,1)
```
Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?


```{r}
vax %>% filter(as_of_date == "2022-02-22") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```

Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="orange") +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rate across California",
       subtitle="population above 36144") +
  geom_hline(yintercept = mean(vax.36$percent_of_population_fully_vaccinated),
             linetype=2,
             color="red")
```

